---
output:
  pdf_document: default
header-includes:
  - \usepackage{color}
---
----
 Summer 2021: Introduction to Data Science
---

```{r}

```


#\textcolor{blue}{Lab 1: Exploratory Data Analysis}
####Author:_____Ian Davis_____

#### \textcolor{red}{Total Grade for Lab 1: /25} 
#### \textcolor{red}{Comments (optional)} 

By now, you have finished or are in the middle of working on the Exploratory Data Analysis sections of the DataCamp course, Introduction to R. This lab helps you to practice what you have learned to our Class Survey dataset. 


## \textcolor{blue}{Template for lab report}
\textbf{Instructions:} This is the template you will use to type up your responses to the exercises. To produce a document that you can print out and turn in just click on Knit PDF above. All you need to do to complete the lab is to type up your BRIEF answers and the R code (when necessary) in the spaces provided below. If you want to find out more about the markdown language click on the MD icon.

It is strongly recommended that you knit your document regularly (minimally after answering each exercise) for two reasons. 

  1. Ensure that there are no errors in your code that would prevent the document from knitting.
  2. View the instructions and your answers in a more legible, attractive format.


## \textcolor{blue}{Load Class Survey data:}

First, download the Class Survey dataset onto your laptop. Then, make sure you upload it to your RStudio workspace on RStudio Server. Refer to the labs from Unit 2 if you need a reminder on how to do this.


-----
\textcolor{blue}{Beginning of Lab 1 Excercises:}
---

#### \textbf{\textcolor{blue}{Exercise 1: }}

Mention two ways of dealing with missing values in your dataset. Compare them to each other in terms of ease of implementation and benefit to your end goal from the project.

#### \textcolor{blue}{Answer 1: } 

Two different ways of dealing with missing values are filtering out rows with missing values prior to applying functions
and using the na.rm argument to filter out missing values as each function is called. The former option has the advantage
of creating a uniform data frame which can be easily manipulated and analyzed (though it may substantially shrink the dataset). The latter option has the advantage of preserving information from rows with missing values though it also makes applying functions to the data frame as a whole (or multiple rows/columns) far more difficult (as the number of elements in each row or column is effectively different). The size of the data frame, the number and location of the missing values, whether functions need to be applied to all/multiple rows and columns, whether the missing values themselves are telling, and the goals of the project all have to be taken into account to determine which method of dealing with missing values is more appropriate.




####\textcolor{red}{Grade for Exercise 1: /5} 
#### \textcolor{red}{Comments:} 



#### \textbf{\textcolor{blue}{Exercise 2: }} 

Using the class survey data provide some descriptive statistics on the variables which correspond to the questions: 

(a)	How many languages do you speak?(Variable: \textit{\textcolor{teal}{Languages}})
(b) How many pets do you own? (This question might not appear in such a format in the dataset.        Construct a new variable that carries this information from the variables you already have. Call    it \textit{\textcolor{teal}{Pets}})
(c) How many siblings do you have? (Variable: \textit{\textcolor{teal}{Siblings}})
(d)	Which school do you attend? (Variable: \textit{\textcolor{teal}{School}})
(e) What color are your eyes? (Variable: \textit{\textcolor{teal}{EyeColor}})
(f) What is your favorite web browser? (Variable: \textit{\textcolor{teal}{WebBrowser}})
(g) How many countries have you visited? (Variable: \textit{\textcolor{teal}{CountriesVisitied}})
(h) What is your favorite season? (Variable: \textit{\textcolor{teal}{FavSeason}})
(i) How many universities/colleges did you apply to in your high school senior year? (Variable: \textit{\textcolor{teal}{CollegeApps}})

#### \textcolor{blue}{Hint: Refer to the first unit labs if you need to remind yourself of how to construct a vector and populate it with data.I believe that this was discussed in Lab 1 in the \textit{Basic Programming} unit}



```{r, eval = TRUE}
##insert the correct commands here (they should not be preceded by a hashtag)
##Exception: Only put a hashtag before the command if it produces an "Error" output.
##    This will ensure that the document knits properly even though the command does not.
##    The graders will still be able to see that you made progress on this question.

setwd("~/LACOL Intro Data Sci/Unit_3")
ClassSurveyData <- read.csv("ClassSurvey.csv")
library(dplyr)

#Languages
languages_vector <- ClassSurveyData$Languages
summary(languages_vector)

#Pets
pets_vector <- ClassSurveyData$Cats + ClassSurveyData$Dogs
ClassSurveyData <- ClassSurveyData %>% mutate(Pets = pets_vector)
summary(pets_vector)

#Siblings
siblings_vector <- ClassSurveyData$Siblings
summary(siblings_vector)

#School
school_vector <- ClassSurveyData$LACOLschool
school_sum <- as.data.frame((summary(school_vector)))
names(school_sum) <- "Number"
school_sum <- school_sum %>% mutate(Percentage = (Number / sum(Number))*100) %>%
                             arrange(desc(Number))
school_sum

#Eye color
eyes_vector <- ClassSurveyData$EyeColor
eyes_sum <- as.data.frame((summary(eyes_vector)))
names(eyes_sum) <- "Number"
eyes_sum <- eyes_sum %>% mutate(Percentage = (Number / sum(Number))*100) %>%
                             arrange(desc(Number))
eyes_sum

#Favorite web browser
web_vector <- ClassSurveyData$WebBrowser
web_sum <- as.data.frame((summary(web_vector)))
names(web_sum) <- "Number"
web_sum <- web_sum %>% mutate(Percentage = (Number / sum(Number))*100) %>%
                             arrange(desc(Number))
web_sum

#Number of countries visited
countries_vector <- ClassSurveyData$CountriesVisited
summary(countries_vector)

#Favorite season
season_vector <- ClassSurveyData$FavSeason
season_sum <- as.data.frame((summary(season_vector)))
names(season_sum) <- "Number"
season_sum <- season_sum %>% mutate(Percentage = (Number / sum(Number))*100) %>%
                             arrange(desc(Number))
season_sum

#College applications
apps_vector <- ClassSurveyData$CollegeApps
summary(apps_vector)
```

#### \textcolor{blue}{Answer 2: } (Make sure that the results of the commands you enteed above are clear and readable. If you suspect that they are not possible to read then copy them here in an organzied manner. Also, add any comments you want/need to elaborate any aspect of the results above)


####\textcolor{red}{Grade for Exercise 2: /10} 
#### \textcolor{red}{Comments:} 



#### \textcolor{blue}{Exercise 3: } 

Use what you have learned in this unit so far to examine the correlation between the following varibales:

(a) Visualize the correlation between the variables \textit{\textcolor{teal}{Pets}} and \textit{\textcolor{teal}{EyeColor}}. Is there any observable trends? If there are trends, do you believe there are any substantive (theoreically sound) reasons for such correlations?

(b) Visualize the correlation between the variables \textit{\textcolor{teal}{Languages}} and \textit{\textcolor{teal}{Countries}}. Is there any observable trends? If there are trends, do you believe there are any substantive (theoreically sound) reasons for such correlations?

(c) Visualize the correlation between the variables \textit{\textcolor{teal}{Countries}} and \textit{\textcolor{teal}{Siblings}}. Is there any observable trends? If there are trends, do you believe there are any substantive (theoreically sound) reasons for such correlations?

####\textcolor{blue}{Make sure to provide the code you used for the visualization and the graphs produced in the following code chunck}




```{r, eval = TRUE}
##insert the correct commands here (they should not be preceded by a hashtag)
##Exception: Only put a hashtag before the command if it produces an "Error" output.
##    This will ensure that the document knits properly even though the command does not.
##    The graders will still be able to see that you made progress on this question.
library(ggplot2)

#pets and eye color
df_pets_eyes <- as.data.frame(pets_vector) 
names(df_pets_eyes) <- "Pets"
df_pets_eyes <- df_pets_eyes %>% mutate(Eye_Color = eyes_vector) %>%
                 group_by(Eye_Color) %>%
                 summarize(Pets_to_People = (mean(Pets, na.rm = TRUE)))

plot_pets_eyes1 <- ggplot(data = df_pets_eyes, mapping = aes(x = Eye_Color, y = Pets_to_People, fill = Eye_Color)) +
                  geom_col()
plot_pets_eyes1


plot_pets_eyes2 <- ggplot(mapping = aes(x = eyes_vector, y = pets_vector)) +
                  geom_count() +
                  scale_size_area(max_size = 10)
plot_pets_eyes2

lm_pets_eyes <- lm(pets_vector ~ eyes_vector)
summary(lm_pets_eyes)

#Languages spoken and countries visited
plot_lang_travel1 <- ggplot(mapping = aes(x = languages_vector, y = countries_vector)) +
                      geom_smooth() +
                      geom_count(alpha = 0.5) +
                      scale_size_area()
plot_lang_travel1

df_lang_countries <- as.data.frame(languages_vector)
names(df_lang_countries) <- "Languages"
df_lang_countries <- df_lang_countries %>% mutate(Countries = countries_vector)
lm_lang_travel <- lm(Languages~ Countries, data = df_lang_countries)
summary(lm_lang_travel)

#Siblings and countries visited

plot_sib_travel <- ggplot(mapping = aes(x = siblings_vector, y = countries_vector)) +
                    geom_count() +
                    geom_smooth()

plot_sib_travel

df_sib_countries <- as.data.frame(siblings_vector)
names(df_sib_countries) <- "siblings"
df_sib_countries <- df_sib_countries %>% mutate(Countries = countries_vector)
lm_sib_travel <- lm(siblings~ Countries, data = df_sib_countries)
summary(lm_sib_travel)

```

#### \textcolor{blue}{Answer 3: } (Provide your answer here regarding each of the correlations examined above.)

There is a visible correlation between eye color and pet ownership as shown in the bar plot of mean pet ownership of people of each eye color shown above. On average, people with blue eyes are more likely to have more pets than people with other eye colors. Furthermore, a linear model reveals the relationship is statistically significant. This is likely largely due to the small proportion of the sample people with blue eyes make up (and, consequently, the ease with which their pet ownership can be extreme). There may be some chance that people with blue eyes are more likely to be White and that Whiteness is connected to pet ownership (e.g. by means of economic privilege to care for a pet). More likely, however, this observed trend is due to the aforementioned make-up of the sample.

With respect to languages spoken and countries visited, a linear plot using geom_smooth and geom_count does not suggest any obvious relationship. While there may be a positive relationship between countries visited and languages spoken at the upper extreme of the relationship, this is only visible in a very small minority of respondents. This lack of an obvious visual relationship is surprising given that one would expect people who travel more to also speak more languages (due to exposure and concerted efforts to learn a language to be able to travel). In fact, a linear model suggests a significant positive effect of the number of languages one speaks on the number of countries one has visited. The visual plot may be misleading.

Concerning one's number of siblings and countries visited, a linear plot using geom_smooth does appear to indicate a negative relationship (particularly if one ignores the one respondent with six siblings who has traveled to several countries). A linear model supports this: more siblings significantly predicts having visited fewer countries. This could be due to the cost of traveling as a large family. It would be interesting to see if the effect disappears if one excludes travel during childhood (perhaps in an older sample).


#### \textcolor{red}{Grade for Exercise 3: /10} 
#### \textcolor{red}{Comments:}





